---
title: "gtsummary in quarto"
format: html
execute: 
  echo: true
  eval: true
---

```{r}
#| message: false
#| echo: true
library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes")))
```

# Customization of `tbl_summary()`

```{r}
#| label: tbl-sex
#| tbl-cap: "Descriptive Statistics Stratified by Sex"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

```{r}
total_sample <- nrow(nlsy)
```

There were `r total_sample` participants.

```{r}
stats <- list(n = nrow(nlsy),
              mean_age = mean(nlsy$age_bir))
```

There were `r stats$n` participants with a mean age of `r stats$nmean_age`.

```{r}
library(gtsummary)
income_table <- tbl_uvregression(
  nlsy,
  y = income,
  include = c(
    sex_cat, race_eth_cat,
    eyesight_cat, income, age_bir
  ),
  method = lm
)
inline_text(income_table, variable = "age_bir")
```

```{r}
#| label: tbl-descr
#| tbl-cap: "Descriptive statistics"
#| output-location: slide
table1 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir)) |> 
  add_overall(last = TRUE)
table1
inline_text(table1, variable = "age_bir", column = "stat_1")
inline_text(table1, variable = "glasses", column = "stat_0",
            pattern = "{n}/{N} ({p}%)")
```

```{r}
inline_text(income_table, variable = "age_bir")
```

```{r}
inline_text(income_table, variable = "age_bir",
            pattern = "{estimate} ({conf.low}, {conf.high})")
```

```{r}
glasses_f <- inline_text(table1, variable = "glasses", 
                         column = "stat_2")
glasses_m <- inline_text(table1, variable = "glasses", 
                         column = "stat_1")
```

A greater proportion of female (`r glasses_f`) than male (`r glasses_m`) participants wore glasses.

```{r}
#| label: tbl-eyesight
#| tbl-cap: "Descriptive Statistics Stratified by Eyesight"
tbl_summary(
  nlsy,
  by = eyesight_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

The descriptive statistics are show stratified by sex (@tbl-sex) and by eyesight (@tbl-eyesight).

```{r}
mean_sleep_weekend <- mean(nlsy$sleep_wknd, na.rm = TRUE)

mean_sleep_weekend
```

People slept an average of `r mean_sleep_weekend` hours.

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```
